{\fontfamily{ptm}\selectfont Lorum ipsum}

### 2025-07-01 Notebook LaTEX/SAGD Under Interpolation/main.tex
>w<, changes to notes

% ==============================================================================
\section{Nesterov's Accelerated Gradient}
    \subsection{In preparations}
        \begin{assumption}[Smooth add nonsmooth]\label{ass:smooth_plus_nonsmooth}
            The function $F = f + g$ where $f:\RR^n \rightarrow \RR$ is a $L$ Lipschitz smooth and $\mu \ge 0$ strongly convex function. 
            The function $g:\RR^n \rightarrow \overline \RR$ is a closed convex proper function. 
        \end{assumption}

        \begin{definition}[Proximal gradient operator]
            Suppose $F = f + g$ satisfies Assumption \ref{ass:smooth_plus_nonsmooth}. 
            Let $\beta > 0$. 
            Then, we define the proximal gradient operator $T_{\beta}$ as 
            \begin{align*}
                T_\beta (x | F) &= \argmin{z} \left\lbrace
                    g(z) + f(x) + \langle \nabla f(x), z - x\rangle + \frac{\beta}{2}\Vert z - x\Vert^2
                \right\rbrace. 
            \end{align*}
        \end{definition}
        \begin{remark}
            If the function $g \equiv 0$, then it yields the gradient descent operator $T_\beta(x) = x - \beta^{-1}\nabla f(x)$. 
            In the context where it's clear what the function $F = f + g$ is, we simply write $T_\beta(x)$ for short. 
        \end{remark}

@@

        \begin{definition}[similar triangle representation of NAPG]\;\label{def:st-method}\\
            Let $(\alpha_k)_{k \ge 0}$ be an R-WAPG sequence. 
            Suppose that the base case $v_{-1}, x_{k - 1}\in \RR^n$ is given to initialize the algorithm. 
            Then the algorithm produces the sequence of iterates $(y_k, x_k, v_k)_{k \ge 0}$ and auxilary parameter sequence $L_k, \tau_k$ satisfying these inequalities: 
            \begin{align*}
                & \tau_k = L_k(1 - \alpha_k)(L_k\alpha_k - \mu)^{-1}, 
                \\
                & y_k = (1 + \tau_k)^{-1} v_{k - 1} + \tau_k (1 + \tau_k)^{-1} x_{k - 1},
                \\
                & D_f(x_k, y_k) \le L_k/2\Vert x_k - y_k\Vert^2, 
                \\
                & x_k = T_{L_k}(y_k),
                \\
                & v_k = x_{k - 1} + \alpha_k^{-1}(x_k - x_{k - 1}). 
            \end{align*} 
        \end{definition}
        The following theorems are critical in analyzing the behavior of algorithm in Definition \ref{def:st-method}. 
        \begin{theorem}[proximal gradient inequality]\label{thm:pg-ineq}
            Let function $F$ satisfies Assumption \ref{ass:smooth_plus_nonsmooth}, so it's $\mu \ge 0$ strongly convex. 
            For all $x\in \RR^n$, define $x^+ = T_L(x)$, then there exists a $B \ge 0$ such that $D_f(x^+, x) \le B/2 \Vert x^+ - x\Vert^2$. 
            Then, for all $z \in \RR^n$ it satisfies proximal gradient inequality at point $x$:  
            \begin{align*}
                0&\le F(z) - F(x^+) - \frac{B}{2}\Vert z - x^+\Vert^2  + \frac{B - \mu}{2}\Vert z - x\Vert^2
                \\
                &=  F(z) - F(x^+) - \langle B(x - x^+), z - x\rangle
                - \frac{\mu}{2}\Vert z - x\Vert^2
                - \frac{B}{2}\Vert x - x^+\Vert^2. 
            \end{align*}
            Since $f$ is assumed to be $L$ Lispchitz smooth, the above condition is true for all $x, y \in \RR^n$ for all $B \ge L$. 
        \end{theorem}
        \begin{remark}
            The theorem is the same as in Nesterov's book \cite[Theorem 2.2.13]{nesterov_lectures_2018}, but with the use of proximal gradient mapping and proximal gradient instead of project gradient hence making it equivalent to the theorem in Beck's book \cite[Theorem 10.16]{beck_first-order_2017}. 
            The only generalization here is parameter $B$ which made to accomodate algorithm that implements Definition \ref{def:st-method} with some line search routine. 
        \end{remark}
        \begin{theorem}[Jensen's inequality]\label{thm:jesen}
            Let $F: \RR^n \rightarrow \overline \RR$ be a $\mu \ge 0$ strongly convex function. 
            Then, it is equivalent to the following condition. 
            For all $x, y \in \RR^n$, $\lambda \in (0, 1)$ it satisfies the inequality 
            \begin{align*}
                (\forall \lambda \in [0, 1])\; 
                F(\lambda x + (1 - \lambda)y) \le \lambda F(x) + (1 - \lambda)F(y) -\frac{\mu\lambda(1 - \lambda)}{2} \Vert y - x\Vert^2. 
            \end{align*}
        \end{theorem}
        \begin{remark}
            If $x, y$ is out of $\dom F$, the inequality still work by convexity. 
        \end{remark}

@@

            Let $F = f + g$ satisfies Assumption \ref{ass:smooth_plus_nonsmooth} for some $L > \mu \ge 0$. 

@@

                \item [(a)] Theorem \ref{thm:pg-ineq}, with $z = z_k$, $k \ge 0$. We can use it because $F = f + g$ satisfies Assumption \ref{ass:smooth_plus_nonsmooth} . 

@@

            Define $F := g + (1/n)\sum_{i = 1}^{n} f_i$, assume that $f_i:\RR^n \rightarrow  \RR$ are all $K_i$ smooth and $\mu^{(i)} \ge 0$ strongly convex function and $g:\RR^n \rightarrow \overline \RR$ is a closed convex proper function. 
            Consequently, the function $f$ cane be written as $F = g + f$ where $f = (1/n)\sum_{i = 1}^{n} f_i$ and it satisfies Assumption \ref{ass:smooth_plus_nonsmooth} with $L = \max_{i = 1, \ldots, n}K_i$ and $\mu = (1/n)\sum_{i = 1}^{n}\mu^{(i)}$. 

@@

        \begin{definition}[SNAPG]\label{def:snapg}

        \end{definition}

>>>>

% ==============================================================================
\section{Nesterov's Accelerated Gradient}
    \subsection{In preparations}
        \begin{assumption}[smooth add nonsmooth]\label{ass:smooth-plus-nonsmooth}
            The function $F = f + g$ where $f:\RR^n \rightarrow \RR$ is a $L$ Lipschitz smooth and $\mu \ge 0$ strongly convex function. 
            The function $g:\RR^n \rightarrow \overline \RR$ is a closed convex proper function. 
        \end{assumption}
        \begin{assumption}[admitting minimizers]\label{ass:smooth-plus-nonsmooth-x}
            Let $F = f + g$ and in addition assume that the set of minimizers $X^+ := \argmin_{x}F(x)$ is non-empty. 
        \end{assumption}

        \begin{definition}[Proximal gradient operator]
            Suppose $F = f + g$ satisfies Assumption \ref{ass:smooth-plus-nonsmooth}. 
            Let $\beta > 0$. 
            Then, we define the proximal gradient operator $T_{\beta}$ as 
            \begin{align*}
                T_\beta (x | F) &= \argmin{z} \left\lbrace
                    g(z) + f(x) + \langle \nabla f(x), z - x\rangle + \frac{\beta}{2}\Vert z - x\Vert^2
                \right\rbrace. 
            \end{align*}
        \end{definition}
        \begin{remark}
            If the function $g \equiv 0$, then it yields the gradient descent operator $T_\beta(x) = x - \beta^{-1}\nabla f(x)$. 
            In the context where it's clear what the function $F = f + g$ is, we simply write $T_\beta(x)$ for short. 
        \end{remark}

@@

        \begin{definition}[similar triangle representation of NAPG]\;\label{def:st-method}\\
            Let $(\alpha_k)_{k \ge 0}$ be an R-WAPG sequence. 
            Suppose that the base case $v_{-1}, x_{k - 1}\in \RR^n$ is given to initialize the algorithm. 
            Then the algorithm produces the sequence of iterates $(y_k, x_k, v_k)_{k \ge 0}$ and auxilary parameter sequence $L_k, \tau_k$ satisfying these inequalities: 
            \begin{align*}
                & \tau_k = L_k(1 - \alpha_k)(L_k\alpha_k - \mu)^{-1}, 
                \\
                & y_k = (1 + \tau_k)^{-1} v_{k - 1} + \tau_k (1 + \tau_k)^{-1} x_{k - 1},
                \\
                & D_f(x_k, y_k) \le L_k/2\Vert x_k - y_k\Vert^2, 
                \\
                & x_k = T_{L_k}(y_k),
                \\
                & v_k = x_{k - 1} + \alpha_k^{-1}(x_k - x_{k - 1}). 
            \end{align*} 
        \end{definition}
        The following theorems are critical in analyzing the behavior of algorithm in Definition \ref{def:st-method}. 
        \begin{theorem}[proximal gradient inequality]\label{thm:pg-ineq}
            Let function $F$ satisfies Assumption \ref{ass:smooth-plus-nonsmooth}, so it's $\mu \ge 0$ strongly convex. 
            For all $x\in \RR^n$, define $x^+ = T_L(x)$, then there exists a $B \ge 0$ such that $D_f(x^+, x) \le B/2 \Vert x^+ - x\Vert^2$. 
            Then, for all $z \in \RR^n$ it satisfies proximal gradient inequality at point $x$:  
            \begin{align*}
                0&\le F(z) - F(x^+) - \frac{B}{2}\Vert z - x^+\Vert^2  + \frac{B - \mu}{2}\Vert z - x\Vert^2
                \\
                &=  F(z) - F(x^+) - \langle B(x - x^+), z - x\rangle
                - \frac{\mu}{2}\Vert z - x\Vert^2
                - \frac{B}{2}\Vert x - x^+\Vert^2. 
            \end{align*}
            Since $f$ is assumed to be $L$ Lispchitz smooth, the above condition is true for all $x, y \in \RR^n$ for all $B \ge L$. 
        \end{theorem}
        \begin{remark}
            The theorem is the same as in Nesterov's book \cite[Theorem 2.2.13]{nesterov_lectures_2018}, but with the use of proximal gradient mapping and proximal gradient instead of project gradient hence making it equivalent to the theorem in Beck's book \cite[Theorem 10.16]{beck_first-order_2017}. 
            The only generalization here is parameter $B$ which made to accomodate algorithm that implements Definition \ref{def:st-method} with some line search routine. 
        \end{remark}
        \begin{theorem}[Jensen's inequality]\label{thm:jesen}
            Let $F: \RR^n \rightarrow \overline \RR$ be a $\mu \ge 0$ strongly convex function. 
            Then, it is equivalent to the following condition. 
            For all $x, y \in \RR^n$, $\lambda \in (0, 1)$ it satisfies the inequality 
            \begin{align*}
                (\forall \lambda \in [0, 1])\; 
                F(\lambda x + (1 - \lambda)y) \le \lambda F(x) + (1 - \lambda)F(y) -\frac{\mu\lambda(1 - \lambda)}{2} \Vert y - x\Vert^2. 
            \end{align*}
        \end{theorem}
        \begin{remark}
            If $x, y$ is out of $\dom F$, the inequality still work by convexity. 
        \end{remark}

@@

            Let $F = f + g$ satisfies Assumption \ref{ass:smooth-plus-nonsmooth} for some $L > \mu \ge 0$. 

@@

                \item [(a)] Theorem \ref{thm:pg-ineq}, with $z = z_k$, $k \ge 0$. We can use it because $F = f + g$ satisfies Assumption \ref{ass:smooth-plus-nonsmooth} . 

@@

            Define $F := g + (1/n)\sum_{i = 1}^{n} f_i$, assume that $f_i:\RR^n \rightarrow  \RR$ are all $K_i$ smooth and $\mu^{(i)} \ge 0$ strongly convex function such that $K_i > \mu^{(i)}$ and, $g:\RR^n \rightarrow \overline \RR$ is a closed convex proper function. 
            Consequently, the function $f$ cane be written as $F = g + f$ where $f = (1/n)\sum_{i = 1}^{n} f_i$ and it satisfies Assumption \ref{ass:smooth-plus-nonsmooth} with $L = \max_{i = 1, \ldots, n}K_i$ and $\mu = (1/n)\sum_{i = 1}^{n}\mu^{(i)}$. 

@@

        This assumption is stronger than Assumption \ref{ass:smooth-plus-nonsmooth}. 
        The interpolation hypothesis from Machine Learning stated that the model has the capacity to perfect fit all the observed data. 
        \begin{assumption}[interpolation hypothesis]\label{ass:interp-hypothesis}
            Suppose that $F := f + (1/n)\sum_{i = 1}^{n} f_i$ satsifying Assumption \ref{ass:sum-of-many}. 
            In addition, assuming that it has $0 = \inf_{x}F(x)$ and, there exists some $\bar x \in \RR^n$ such that for all $i = 1, \ldots, n$ it satisfies $0 = f_i(\bar x)$. 
            Obviously, all such $\bar x$ forms the set of minimizers of $F$. 

        \begin{definition}[SNAPG]\label{def:snapg}
            Let $F$ satisfies Assumption \ref{ass:sum-of-many}. 
            Let $(I_k)_{k \ge 0}$ be a list of i.i.d random variables uniformly sampled from set $\{0, 1, 2, \cdots, n\}$. 
            Initialize $v_{-1} = x_{-1}, \alpha_0 = 1$. 
            The SAGD generates the sequence $(y_k, x_k, v_k)_{k \ge 0}$ such that for all $k \ge 0$ they satisfy: 
            \begin{align*}
                & L_k/L_{k - 1}(1 - \alpha_{k})\alpha_{k - 1}^2 = \alpha_{k}\left(\alpha_{k} - \mu^{(I_k)}/L_k\right), \\
                & \tau_k = L_k(1 - \alpha_k)\left(L_\alpha \alpha_k - \mu^{(I_k)}\right)^{-1}, \\
                & y_k = (1 + \tau_k)^{-1}v_{k - 1} + \tau_k(1 + \tau_k)^{-1}x_{k - 1}, \\
                & x_k =  T_{L_k}(y_k | F_{I_k}) \text{ s.t: } D_f(x_k, y_k) \le L_k/2\Vert y_k - x_k\Vert^2, \\
                & v_k = x_{k - 1} + \alpha_k^{-1}(x_k - x_{k - 1}). 
            \end{align*}
        \end{definition}
        \begin{remark}
            The sequence $\alpha_k$ is a random variable because it changes according to variable $I_k$ at the current iteration.
            This decision is made because if Theorem \ref{thm:pg-ineq} needs to hold during each iteration of the algorithm.
        \end{remark}

        \begin{theorem}[SNAPG one step inequality]\label{thm:snapg-one-step}
